04-plot characteristics

### 2018-05-15 01_writing/Schratz_hyperspectral.tex
15/05/2018

% link to PDF with veg indeces
All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that are offered by the \texttt{hsdar} package have been calculated.
Additionally, all possible \ac{NRI} were calculated from the data using the formula:

\begin{equation}
	NRI_{i,j} = \frac{B_{i} - B_{j}}{B_{i} + B_{j}}
\end{equation}

@@

@@

After the plot level analysis, we merged all observations into a single dataset and fitted another \ac{RR} model.
We used a block-level based spatial \ac{CV} for this setup.
For a total number of $m$ folds ($n$ = number of plots), every plot is once the test set while the training is done on the remaining ones.
This ensures that the test set is fully spatially independent from the training set.
Tuning was also performned on the block level within the training set on $n - 1$ folds using 200 random search iterations.
Due to the fixing of the indices on a plot level, varying those to use multiple repetitions is not possible.

@@

\subsection{Predictive performances}

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{Predictive performances of lasso, ridge and elasticnet penalties on the single plots and the merged dataset using \ac{RMSE} as the error measure.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Penalty & Lasso  & Ridge  & Elasticnet \\
	\hline
	Laukiz 1     & 146.33 & 89.89  & 135.58     \\
	Laukiz 2     & 172.70 & 30.37  & 37.07      \\
	Luiando      & 74.85  & 76.02  & 74.77      \\
	Oiartzun     & 327.93 & 106.65 & 260.38     \\
	All Plots    & 56.88  & 55.88  & 56.76      \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:penalty_comparison}
\end{table}

\begin{table}[b!]
\centering
\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.}
\begingroup\footnotesize
\begin{tabular}{lll}
	\\
	Plot     & Supermodel (Block CV) & Single plot (SpCV) \\
	\hline
	Laukiz 1 & 58.95                 & 89.89              \\
	Laukiz 2 & 27.94                 & 30.37              \\
	Luiando  & 69.72                 & 76.02              \\
	Oiartzun & 58.09                 & 106.65             \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}

>>>>

% link to PDF with veg indeces
All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data and offered by the \texttt{hsdar} package have been calculated.
Additionally, all possible \ac{NRI} were calculated from the data using the formula:

\begin{equation}
	NRI_{i,j} = \frac{b_{i} - b_{j}}{b_{i} + b_{j}}
\end{equation}

@@

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{10-fold 20-times repeated spatial \ac{CV} performances of lasso, ridge and elasticnet penalties on the plot level and the merged dataset using \ac{RMSE} as the error measure. Values show the overall mean and standard deviation at the repetition level.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Penalty   & Lasso         & Ridge        & Elasticnet    \\
	\hline
	Laukiz 1       & 133.15 (4.66) & 86.71 (2.96) & 121.04 (4.15) \\
	Laukiz 2       & 91.93 (6.78)  & 29.73 (0.38) & 47.18 (13.70) \\
	Luiando        & 74.85         & 76.02        & 74.77         \\
	Oiartzun       & 327.93        & 106.65       & 260.38        \\
	Merged dataset & 56.88         & 55.88        & 56.76         \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:penalty_comparison}
\end{table}

@@

After the plot level analysis, we merged all observations into a single dataset and fitted another \ac{RR} model.
We used a block-level based spatial \ac{CV} for this setup.
For a total number of $m$ folds ($n$ = number of plots), every plot is once the test set while the training is done on the remaining ones.
This ensures that the test set is fully spatially independent from the training set.
Tuning was also performned on the block level within the training set on $n - 1$ folds using 200 random search iterations.
Due to the fixing of the indices on a plot level, varying those to use multiple repetitions is not possible.
The idea behind fitting a supermodel is that this model includes information from all plots rather than just from a single plot.
This may possibly reduce the predictive error and create a more robust model.

@@

\subsection{Plot characteristics}

Luiando shows the highest defoliation ($\bar{x} = 68.36 \%$) among the plots while Laukiz 2 is the healthiest ($\bar{x} = 17.73 \%$) (\autoref{fig:defol_boxplots}).
Laukiz 1 and Oiartzun both show a medium defoliation level around 50 \%.
Oiartzun consists mainly of trees that show either a very high level of defoliation ($70 \% <=$) or none at all.
Laukiz 1 shows an evenly distributed level of defoliation across the entire plot.

The high defoliation level of Luiando and Oiartzun is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
Laukiz 2 shows a mean tree density of 62.51 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.35 (Laukiz 1), 33.77 (Luiando), 35.02 (Oiartzun)) (\autoref{fig:rmse_defol_dens}).

% defoliation boxplots
\begin{figure} [t!]
	\begin{center}
		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {../03_figures/results/boxplot_defol.pdf}}
		\caption{Descriptive statistics of the response variable \textit{defoliation}.}
		\label{fig:defol_boxplots}
	\end{center}
\end{figure}

% spectral signatures
\begin{figure} [t!]
	\begin{center}
		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/spectral_signatures.pdf}}
		\caption{Spectral signatures (mean and standard deviation) of each plot.
			The colored lines show the most important band for each plot, respectively: Band 80 (773nm, red), band 89 (817nm, blue), band 95 (846nm, orange), band 23 (503nm, green).}
		\label{fig:spectral_signatures}
	\end{center}
\end{figure}

\subsection{Predictive performance}

\begin{table}[t!]
\centering
\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure. The values for "merged dataset" correspond to the fold for which the respective plot was serving as the test set. For "single plot", the values correspond to the mean value of the SpCV at the repetition level (10 folds, 20 repetitions).}
\begingroup\footnotesize
\begin{tabular}{lll}
	\\
	Plot/Data & Merged dataset (Block CV) & Single plot (SpCV) \\
	\hline
	Laukiz 1  & 58.95                     & 89.89              \\
	Laukiz 2  & 27.94                     & 30.37              \\
	Luiando   & 69.72                     & 76.02              \\
	Oiartzun  & 58.09                     & 106.65             \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}

\ac{RR} shows the lowest error for three out of four plots (for Luiando \textit{elasticnet} shows a slightly better performance) (\autoref{tab:penalty_comparison}).
The magnitude of difference for \ac{RR} compared to the other penalties for the plots in which \ac{RR} showed the best performance ranges between XX and XX percent.
For the merged dataset, all penalties show a similar mean predictive performance that outperform all single plot models besides the Laukiz 2 model.

When comparing the mean predictive performance of the plot level model against the performance of the super model at the plot level (when the respective plot served as the test set), the supermodel also outperforms the Laukiz 2 model (27.94 vs 30.37 RMSE) (\autoref{tab:supermodel_performance}).

The worst performance of the supermodel on the fold level is reported for Luiando (69.72 RMSE) while for the single plot models Oiartzun shows the highest error (106.65 RMSE).

% rmse vs defol and point dens
\begin{figure} [b!]
	\begin{center}
		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/mean-defol_mean-point-dens.pdf}}
		\caption{RMSE vs. mean defoliation and point density}
		\label{fig:rmse_defol_dens}
	\end{center}
\end{figure}

### 2018-05-15 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
15/05/2018

@@

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
```

>>>>

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge_bf2_10folds.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

@@

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
```

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

### 2018-05-15 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
15/05/2018

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
```

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
```

>>>>

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
```

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

@@

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
```

Calculate repetition mean

```{r}
resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
resa_glmnet$measures.test %>% 
  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
  group_by(rep) %>% 
  summarise_at("rmse", mean) %>% 
  summarise(mean(rmse), sd(rmse))
```

### 2018-05-15 02_scripts/04_plot_characteristics.Rmd
15/05/2018

library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(raster)
library(ggplot2)
library(hrbrthemes)
library(cowplot)
```

@@

```{r}
data %>%
  group_by(plot) %>%
  ggplot(aes(y = deftot, x = plot)) +
  geom_boxplot() +
  theme_ipsum_rc()
```

@@

```{r}
rmse_dens <- ggplot(df, aes(x = point_dens, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean point density (m)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

@@

```{r}
rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

@@

```{r}
legend <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE", color = "Plot") +
  theme_ipsum_rc()
legend <- get_legend(legend)
```

@@

```{r}
plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
```

>>>>

library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(raster)
library(ggplot2)
library(hrbrthemes)
library(cowplot)
library(here)
library(ggpubr)
library(ggsci)
library(ggrepel)
```

@@

```{r}
boxplot_defol = data %>%
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "deftot", color = "plot",
    add = "jitter", add.params = list(size = "deftot")
  ) +
  scale_size(range = c(0.5, 0.5)) +
  annotate("text",
    label = expression(bold(atop("n = 559", bar(x)~"= 53.23"))), x = 1, 
    y = 112, size = 3, colour = "#BC3C29", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 382", bar(x)~"= 17.73"))), x = 2, 
    y = 112, size = 3, colour = "#0072B5", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 301", bar(x)~"= 68.36"))), x = 3, 
    y = 112, size = 3, colour = "#E18727", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 499", bar(x)~"= 44.64"))), x = 4, 
    y = 112, size = 3, colour = "#20854E", fontface = 2
  ) +
  scale_color_nejm() +
  theme(legend.position = "none") +
  labs(y = "Total defoliation per tree (%)", x = "Plot")
boxplot_defol
```

```{r}
save_plot(here("03_figures/results/boxplot_defol.png"), boxplot_defol)
```

@@

```{r}
rmse_dens <- ggplot(df, aes(x = point_dens, y = rmse)) +
  geom_point(aes(color = plot), shape = 19, size = 3) +
  labs(x = "Mean point density (m)", y = "RMSE") +
  theme_ipsum_rc(
    axis_title_size = 22,
    axis_title_just = "c"
  ) +
  scale_color_nejm() +
  theme(
    plot.margin = unit(c(6, 0, 6, 0), "pt"),
    legend.position = "none",
    axis.text = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.title.y = element_text(margin = margin(
      t = 0, r = 15,
      b = 0, l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 15, r = 0,
      b = 0, l = 0
    ))
  )
```

@@

```{r}
rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19, size = 3) +
  labs(x = "Mean defoliation (%)", y = "RMSE") +
  theme_ipsum_rc(
    axis_title_size = 22,
    axis_title_just = "c"
  ) +
  scale_color_nejm() +
  theme(
    plot.margin = unit(c(6, 0, 6, 0), "pt"),
    legend.position = "none",
    axis.text = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.title.y = element_text(margin = margin(
      t = 0, r = 15,
      b = 0, l = 0
    )),
    axis.title.x = element_text(margin = margin(
      t = 15, r = 0,
      b = 0, l = 0
    ))
  )
```

@@

```{r}
legend <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19, size = 3) +
  labs(x = "Mean defoliation (%)", y = "RMSE", color = "Plot") +
  theme_ipsum_rc(
    axis_title_size = 16,
    axis_title_just = "c"
  ) +
  scale_color_nejm() +
  theme(
    axis.text = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    axis.title.y = element_text(margin = margin(
      t = 0, r = 15,
      b = 0, l = 0
    ))
  )
legend <- get_legend(legend)
```

@@

```{r}
plots_grid <- plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
```

```{r}
save_plot(here("03_figures/results/mean-defol_mean-point-dens.png"), plots_grid, ncol = 3)
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
#        parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

# copy to presentation folder
# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
#           overwrite = TRUE)
```

# Session Info

```{r}
sessionInfo()
```
